Cassini State Capture

### 2019-06-26 initial/notes.tex
stash writeup
stash

@@

@@

Can we analytically show that the Melnikov splitting dominates, and that in
particular the seeming small excess of $P_{hop}$ is predictable?

>>>>

\textbf{NB:} Our convention is that $\phi = 0$ for CS1, 3, 4, all of which are
on opposite sides of $\hat{l}$ from $\hat{l}_p$.

@@

\textbf{Note:} The full formula, by actually evaluating all the terms, comes out
to be
\begin{equation}
    P_{hop} = \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi
        \p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
        + 8\eta^{3/2}\cos I \sqrt{\sin I}}.
\end{equation}
It bears noting this formula slightly overpredicts $P_{hop}$ compared to the
actual simulation values; this comes down to ICs too near the separatrix being
necessarily escape-bound (can see \autoref{fig:CS4s} for more evidence).

@@

\textbf{Revisited:} We should be able to evaluate this integral analytically.
We've already mostly done the $\dot{\mu}\;\mathrm{d}\phi$ integral, let's now
consider the $\frac{\dot{s}}{\dot{\phi}}\Delta H_{,s}\;\mathrm{d}\phi$
component. This is just (we change sign convention, $\Delta H = H - H_4$ now,
per recent developments)
\begin{align*}
    \int_{\mathcal{C}_{\pm}} \frac{\dot{s}}{\dot{\phi}^{(0)}}
            \pd{\Delta H}{s}\;\mathrm{d}\phi &=
        \int \dot{s}\frac{-\frac{\mu^2}{2s_c} + \frac{\cos^2 Is_c}{2s^2}}{
            -\frac{s}{s_c}\mu + \cos I}\;\mathrm{d}\phi,\\
        &= \int \frac{\dot{s}s_c}{2s^2}\s*{\frac{s\mu}{s_c} + \cos
            I}\;\mathrm{d}\phi,\\
        &= \pm \int \epsilon 2\Omega_1
            \p*{\mu_{\pm} - \frac{s}{2\Omega_1}\p*{1 + \mu_{\pm}^2}}
                \frac{s_c}{2s^2} \sqrt{2\frac{s_c}{s} \sin I\p*{1 - \cos \phi}}
                \;\mathrm{d}\phi,\\
        &\approx \pm \frac{\epsilon \Omega_1 s_c}{s^2}
            \int \p*{\mu_{\pm} - \frac{s}{2\Omega_1}}
            \sqrt{2\frac{s_c}{s} \sin I\p*{1 - \cos \phi}} \;\mathrm{d}\phi,\\
        &\approx \pm\frac{\epsilon \Omega_1 s_c}{s^2}
            \int \p*{\mu_4 - \frac{s}{2\Omega_1} \pm
                \sqrt{2\frac{s_c}{s} \sin I\p*{1 - \cos \phi}}}
            \sqrt{2\frac{s_c}{s} \sin I\p*{1 - \cos \phi}} \;\mathrm{d}\phi,\\
        &= \frac{\epsilon \Omega_1 s_c}{s^2}\s*{
            \pm \p*{\mu_4 - \frac{s}{2\Omega_1}}\p*{-8\sqrt{\eta \sin I}}
                + 4\pi \frac{s_c}{s}\sin I}.
\end{align*}
Indeed, the sum over the two legs is positive, which is expected: during
spindown, the separatrix expands, so the separatrix must have $\Delta H > 0$
owing just to adiabatic expansion.

For the $\dot{\mu}$ integral, the integrand is $\p*{\frac{2\Omega}{s} - \mu}$
times the original integrand; this means the final result is $\frac{2\Omega}{s}$
times the old result plus a contribution approximately $-\int \epsilon \mu
\;\mathrm{d}\phi$. This latter term is a purely antisymmetric contribution, so
we may approximate
\begin{align*}
    \int_{\mathcal{C}_{\pm}}\dot{\mu}^{(1)}\;\mathrm{d}\phi
        &\approx \epsilon\frac{2\Omega}{s}\p{\mp 2\pi
            + 16\cos I \eta^{3/2}\sqrt{\sin I}}
            + 8\epsilon\sqrt{\eta \sin I},\\
        &\approx \mp \frac{4\pi\epsilon \Omega}{s} + 8\sqrt{\eta \sin I}.
\end{align*}
This is in agreement with our work from 5.1. Thus, the total expression is

### 2019-06-26 writeups/paper.tex
stash

    \documentclass[
        fleqn,
        usenatbib,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
    }
    \usepackage[T1]{fontenc}
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}

@@

With a dissipitive term that favors alignment, we expect that separatrix
encounters can only occur from zone II and zone III\@. We will focus on the
outcomes of trajectories origiating in zone III for now. We now aim to compute
\autoref{eq:p_c_gen}, for which we need:
\begin{subequations}
    \begin{align}
        \Delta_- &= \int\limits_{\mathcal{C}_-}
                \phi'\epsilon\p{1 - \mu^2}\;\mathrm{d}t,\nonumber\\
            &= \int\limits_{\mathcal{C}_s}
                \p{1 - \mu^2}\;\mathrm{d}\phi,\\
        \Delta_- + \Delta_+ &= \int\limits_{\mathcal{C}_- + \mathcal{C}_+}
            \epsilon\p{1 - \mu^2}\;\mathrm{d}\phi.
    \end{align}
\end{subequations}
Note that along $\mathcal{C}_{\pm}$ that $\mathrm{d}\phi$ has sign $\mp$
respectively, thanks to our sign convention. We proceed to integrate:
\begin{itemize}
    \item $\Delta_-$: Recalling the parameterization for the separatrix
        $\mu_{sep}(\phi)$ given by \autoref{eq:mu_sep}, we can approximate the
        integrand as $1 + \mathcal{O}(\eta^1)$. This yields simply $\Delta_-
        \approx 2\pi \epsilon + \mathcal{O}(\eta^1)$.

    \item $\Delta_- + \Delta_+$: Since $\mathrm{d}\phi$ changes signs from
        $\mathcal{C}_-$ to $\mathcal{C}_+$, only terms in $(1 - \mu^2)$ that
        also change sign over $\mathcal{C}_{\mp}$ contribute to the integral.
        The leading order term that does so can be read off of
        \autoref{eq:mu_sep}, so the integral evaluates as
        \begin{align}
            \Delta_- + \Delta_+ &\approx
                2\int\limits_0^{2\pi}
                    \s{(\eta \cos I) \sqrt{2\eta \sin I(1 - \cos \phi)}
                        + \mathcal{O}(\eta^{5/2})} \;\mathrm{d}\phi,\\
                &\approx 32\epsilon \eta^{3/2} \cos I \sqrt{\sin I}
                    + \mathcal{O}(\eta^{5/2}).
        \end{align}
\end{itemize}
From these two integrals, we obtain
\begin{equation}
    P_c = \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.\label{eq:p1pc}
\end{equation}

As evidence, we provide the following table of separatrix capture probabilities
in \autoref{tab:p_c_num} (TODO make a plot and more robust parameter
exploration):
\begin{table*}
    \centering
    \begin{tabular}{c|c c c c}
        $\eta$ & $0.025$ & $0.05$ & $0.1$ & $0.2$\\\midrule
        Numerical $P_c$ & $0.010 \pm 0.002$ & $0.028 \pm 0.003$
            & $0.085 \pm 0.004$ & $0.251 \pm 0.006$ \\
        Analytical $P_c$ & $0.0111$ & $0.0313$ & $0.0886$ & $0.2503$.
    \end{tabular}
    \caption{Capture probability for four different values of $\eta$, all using
    $\epsilon = 3 \times 10^{-4}$. Different values of $\epsilon$ were tried
    $\epsilon \in 10^{[-2, -4]}$ with no effect on $P_c$. $10000$ random initial
    conditions were used for the numerical calculations, of which roughly $1/2$
    experience separatrix crossing. Numerical uncertainties are estimated as
    $\sqrt{N}$ errors.}\label{tab:p_c_num}
\end{table*}

For trajectories originating in zone II, numerical simulations indicate they can
never leave the separatrix. This can be intuited by considering trajectories
that are near the separatrix; over one orbit, they accumulate $\Delta h =
\Delta_+ + \Delta_- > 0$, which acts to drive them \emph{towards} CS2 and away
from the separatrix. Thus, separatrix crossing is forbidden in zone II, and our
above characterization of separatrix encounters with $P_c$ in \autoref{eq:p1pc}
is complete.

TODO prove the above completely.

@@

\subsection{Spindown Model}

We may first consider the unperturbed Hamiltonian
system with $\rd{\eta}{t} = \frac{\eta}{t_\eta}$ growth with characteristic
timescale $t_\eta$. Again rescaling time $\tau \equiv \alpha t$ and defining
$\epsilon \equiv \frac{1}{\alpha t_\eta}$, we obtain $\eta' = \epsilon
\eta$, where $\epsilon > 0$. Such a model is motivated e.g.\ when a planet's
spin $\hat{s}$ orbits a host star with angular momentum $\hat{l}$ while the star
spins down; $\dot{\alpha} < 0$ here, and $\eta' > 0$.

We then seek $\Delta_{\pm}$, determined by
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}}
            \pd{h}{\eta}\eta'\;\mathrm{d}\tau,\nonumber\\
        &= \int\limits_{\mathcal{C}_{\pm}}
            \frac{\pd{h}{\eta}}{\phi'}\eta'\;\mathrm{d}\phi.
\end{align}
We must proceed carefully since $\phi' \to 0$ at the endpoints of
$\mathcal{C}_{\pm}$; careful work shows these zeros cancel against zeros of
$\pd{h}{\eta}$, and
\begin{equation}
    \Delta_{\pm} = \int\limits_{\mathcal{C}_{\pm}} \eta'
        \p{\cos I \mp \sin I \sqrt{\frac{1 - \cos \phi}{2\eta\sin I}}}
            \;\mathrm{d}\phi.\label{eq:p2_delta}
\end{equation}
Immediately, we see that if we use $\eta' = \epsilon \eta$, then $\Delta_{\pm} >
0$ and \emph{all separatrix encounters result in capture} regardless of whether
the trajectory originates in zone I or zone III\@. Since increasing $\eta$
results in larger separatrix, zone II trajectories will not experience
separatrix crossing. Indeed, an example of such a simulation is presented in
\autoref{fig:2_1testo4}. Further simulations with other initial conditions
support the guaranteed capture prediction in this system.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/1testo4.png}
    \caption{Simulation with initial condition in zone I subject to $\eta' =
    \epsilon \eta$, where $\epsilon = 10^{-4}$. The top panel shows the
    trajectory (red) and separatrix at separatrix crossing (dashed black) in
    the labelled coordinates. The bottom panel shows the evolution of the
    enclosed area by the trajectory (red dots), area of the separatrix (blue)
    and $\eta$ (dotted green) over time.}\label{fig:2_1testo4}
\end{figure}

@@

\subsection{Toy Model}

It is clear that $\eta'$ in \autoref{eq:p2_delta} being the same sign along both
$\mathcal{C}_{\pm}$ is the source of this guaranteed capture. To assess the
accuracy of our formalism, we choose to further study $\eta' = \epsilon \eta
\mu$. Straightforward evaluation of \autoref{eq:p2_delta} yields
\begin{equation}
    \Delta_{\pm} = \pm \epsilon \eta_\star \s{
        2\pi \sin I \pm 12 \sqrt{\eta_\star \sin I}\cos I
            + 2\pi \eta_\star \cos^2 I}.
\end{equation}
It bears noting that we must evaluate these integrals at $\eta = \eta_\star$ the
value at separatrix crossing. Since $\eta$ evolves over time, $\eta_{\star}$
depends on initial conditions. While one might expect the adiabatic invariance
of the action variable determines $\eta_\star$, this turns out to not be the
case; we address this in \autoref{ss:noncons}.

Since $\abs{\Delta_+} > \abs{\Delta_-}$, separatrix encounters originating in
zone III always result in capture. However, encounters originating in zone I
experience capture probability
\begin{equation}
    P_c = \frac{24\cos I \sqrt{\eta_\star \sin I}}{
        2\pi \sin I + 12\sqrt{\eta_\star \sin I}\cos I
            + 2\pi \eta_\star \cos^2 I}.
\end{equation}
This is in good agreement
% TODOYUBO plots

\subsection{Non-conservation of Adiabatic Invariant}\label{ss:noncons}

Best evidence is $\eta_{\star}$ when changing $I$, blatantly wrong scaling

@@

% TODOYUBO show that zone II stays in zone II?

>>>>

    \documentclass[
        fleqn,
        usenatbib,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
    }
    \usepackage[T1]{fontenc}
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures
    \captionsetup{compatibility=false}

@@

With a dissipitive term that favors alignment, we expect that separatrix
encounters can only occur from zone II and zone III\@. We will focus on the
outcomes of trajectories origiating in zone III for now. We now aim to compute
\autoref{eq:p_c_gen}, for which we need:
\begin{subequations}
    \begin{align}
        \Delta_- &= \int\limits_{\mathcal{C}_-}
                \phi'\epsilon\p{1 - \mu^2}\;\mathrm{d}t,\nonumber\\
            &= \int\limits_{\mathcal{C}_s}
                \p{1 - \mu^2}\;\mathrm{d}\phi,\\
        \Delta_- + \Delta_+ &= \int\limits_{\mathcal{C}_- + \mathcal{C}_+}
            \epsilon\p{1 - \mu^2}\;\mathrm{d}\phi.
    \end{align}
\end{subequations}
Note that along $\mathcal{C}_{\pm}$ that $\mathrm{d}\phi$ has sign $\mp$
respectively, thanks to our sign convention. We proceed to integrate:
\begin{itemize}
    \item $\Delta_-$: Recalling the parameterization for the separatrix
        $\mu_{sep}(\phi)$ given by \autoref{eq:mu_sep}, we can evaluate
        explicitly over $\mathcal{C}_-$. This gives result
        \begin{equation}
            \Delta_- = 2\pi\p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
                + 16\cos I \eta^{3/2}\sqrt{\sin I}.
        \end{equation}

    \item $\Delta_- + \Delta_+$: Note $\mathrm{d}\phi$ changes signs between
        $\mathcal{C}_-$ and $\mathcal{C}_+$. Thus, only terms in $\Delta_-$ that
        also change signs on $\mathcal{C}_+$ contribute to the integral. This
        allows us to evaluate
        \begin{align}
            \Delta_- + \Delta_+ &= 32\epsilon \eta^{3/2} \cos I \sqrt{\sin I}.
        \end{align}
\end{itemize}
From these two integrals, we obtain
\begin{equation}
    P_c = \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi
        \p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
        + 8\eta^{3/2}\cos I \sqrt{\sin I}}.\label{eq:p1pc}
\end{equation}

\begin{table*}
    \centering
    \begin{tabular}{c|c c c c}
        $\eta$ & $0.025$ & $0.05$ & $0.1$ & $0.2$\\\midrule
        Numerical $P_c$ & $0.010 \pm 0.002$ & $0.029 \pm 0.003$
            & $0.082 \pm 0.004$ & $0.256 \pm 0.006$ \\
        Analytical $P_c$ & $0.0112$ & $0.0320$ & $0.0915$ & $0.2627$.
    \end{tabular}
    \caption{Capture probability for four different values of $\eta$, all using
    $\epsilon = 3 \times 10^{-4}$. Different values of $\epsilon$ were tried
    $\epsilon \in 10^{[-2, -4]}$ with no effect on $P_c$. $10000$ random initial
    conditions uniformly distributed in $(\mu, \phi)$ were used for the
    numerical calculations, of which roughly $1/2$ experience separatrix
    crossing. Numerical uncertainties are estimated as $\sqrt{N}$ where $N$ is
    the count per bin.}\label{tab:p_c_num}
\end{table*}
As evidence, we provide the following table of separatrix capture probabilities
in \autoref{tab:p_c_num}.

It may be noted that the numerical $P_c$ are systematically low compared to
their analytical estimates. This has a clean explanation: initial conditions
that start too close to the separatrix violate the assumption where $h_i$ is
uniformly distributed on $[-\Delta_-, 0]$. In particular, most of the points
near the separatrix and in zone III are escaping points, as can be visually seen
in \autoref{fig:0_6manifolds}. A systematic underdensity of points converging to
CS2 (separatrix capture) can be seen on the right hand side of the plot near the
separatrix in \autoref{fig:0_3stats}.

A further interesting feature in \autoref{fig:0_3stats} is a small portion of
points in zone II near CS4  at $\phi = 2\pi$ that escape. These points accrue
$\Delta_+$ on their first orbit and escape without librating. Visually, this
zone is obvious in \autoref{fig:0_6manifolds} as the separation between
$W_s^{(1)} \cup W_u^{(1)}$ and $W_s^{(0)} \cup W_u^{(0)}$ near $\phi = 2\pi$.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/3stats3_5_0_1.png}
    \caption{Plot of the CSs that initial conditions converge to. Each subplot
    has a set of initial conditions (black dots) and the CS (red dot) that each
    of these initial conditions converge to. Initial conditions are chosen
    uniformly in $(\mu, \phi)$ coordinates. The parameters for the simulation
    are in the title.}\label{fig:0_3stats}
\end{figure}

For trajectories originating in zone II, numerical simulations indicate they can
never leave the separatrix. This can be understood quantitatively (excepting the
few initial conditions in \autoref{fig:0_3stats} discussed above). For any orbit
in zone II that completes a full libration over contour $\mathcal{C}$, the total
change in $h$ is given
\begin{equation}
    \Delta h = \epsilon \oint_{\mathcal{C}} (1 - \mu^2)\;\mathrm{d}\phi.
\end{equation}
Since $\mathcal{C}$ is a closed contour, any terms that are symmetric on the top
and bottom legs of $\mathcal{C}$ vanish. If we decompose $\mu(\phi) \equiv
\bar{\mu} \pm \mu'(\phi)$ where $\bar{\mu}$ is the average value of $\mu(\phi)$
over $\mathcal{C}$ and $\mu'(\phi) \geq 0$ everywhere, then it is clear
\begin{equation}
    \Delta h = 2\epsilon \int\limits_0^{2\pi} 2\bar{\mu}\mu'(\phi)
            \;\mathrm{d}\phi.
\end{equation}
Now, since $\mu'(\phi) \geq 0$ everywhere, and $\bar{\mu} \in [\mu_2, \mu_4] >
0$ as well, $\Delta h > 0$ necessarily, and \emph{all full librations in zone II
will be driven towards CS2} (the point of maximum $h$ and maximum $H^{(0)}$).

@@

\subsection{Spindown Model}\label{ss:p2_spindown}

We may first consider the unperturbed Hamiltonian
system with $\rd{\eta}{t} = \frac{\eta}{t_\eta}$ growth with characteristic
timescale $t_\eta$. Again rescaling time $\tau \equiv \alpha t$ and defining
$\epsilon \equiv \frac{1}{\alpha t_\eta}$, we obtain
\begin{equation}
    \rd{\eta}{t} = \epsilon \eta,
\end{equation}
where $\epsilon > 0$. Such a model is motivated e.g.\ when a planet's
spin $\hat{s}$ orbits a host star with angular momentum $\hat{l}$ while the star
spins down; $\dot{\alpha} < 0$ here, and $\eta' > 0$.

We then seek $\Delta_{\pm}$, determined by
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}}
            \pd{h}{\eta}\eta'\;\mathrm{d}\tau,\nonumber\\
        &= \int\limits_{\mathcal{C}_{\pm}}
            \frac{\pd{h}{\eta}}{\phi'}\eta'\;\mathrm{d}\phi.
\end{align}
We must proceed carefully since $\phi' \to 0$ at the endpoints of
$\mathcal{C}_{\pm}$; careful work shows these zeros cancel against zeros of
$\pd{h}{\eta}$, and
\begin{equation}
    \Delta_{\pm} = \int\limits_{\mathcal{C}_{\pm}} \eta'
        \p{\cos I \mp \sin I \sqrt{\frac{1 - \cos \phi}{2\eta\sin I}}}
            \;\mathrm{d}\phi.\label{eq:p2_delta}
\end{equation}
Immediately, we see that if we use $\eta' = \epsilon \eta$, then $\Delta_{\pm} >
0$ and \emph{all separatrix encounters result in capture} regardless of whether
the trajectory originates in zone I or zone III\@. Since increasing $\eta$
results in larger separatrix, zone II trajectories will not experience
separatrix crossing. Indeed, an example of such a simulation is presented in
\autoref{fig:2_1testo4}. Further simulations with other initial conditions
support the guaranteed capture prediction in this system.
\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{../initial/2_toy2/1testo4.png}
    \caption{Simulation with initial condition in zone I subject to $\eta' =
    \epsilon \eta$, where $\epsilon = 10^{-4}$. The top panel shows the
    trajectory (red) starting with initial condition $\mu_0 = 0.140, \phi_0 = 0$
    (green dot) and separatrix at separatrix crossing (dashed black) in
    the labelled coordinates. The end of the simulation is labelled in the blue
    cyan dot. The bottom panel shows the evolution of the enclosed area by the
    trajectory (red dots), area of the separatrix (blue) and $\eta$ (dotted
    green) over time.}\label{fig:2_1testo4}
\end{figure}

@@

\subsection{Nontrivial Test Problem}\label{ss:p2_toy}

It is clear that $\eta'$ in \autoref{eq:p2_delta} being the same sign along both
$\mathcal{C}_{\pm}$ is the source of this guaranteed capture. To assess the
accuracy of our formalism, we propose instead test problem where $\eta$ is
varied following
\begin{equation}
    \eta' = \epsilon \eta \mu.
\end{equation}
Straightforward evaluation of \autoref{eq:p2_delta} yields
\begin{equation}
    \Delta_{\pm} = \pm \epsilon \eta_\star \s{
        2\pi \sin I \pm 12 \sqrt{\eta_\star \sin I}\cos I
            + 2\pi \eta_\star \cos^2 I}.
\end{equation}
It bears noting that we must evaluate these integrals at $\eta = \eta_\star$ the
value at separatrix crossing. Since $\eta$ evolves over time, $\eta_{\star}$
depends on initial conditions. While one might expect the adiabatic invariance
of the action variable determines $\eta_\star$, this turns out to not be the
case; we address this in \autoref{ss:noncons}.

Since $\abs{\Delta_+} > \abs{\Delta_-}$, separatrix encounters originating in
zone III always result in capture. However, encounters originating in zone I
experience capture probability
\begin{equation}
    P_c(\eta_\star) = \frac{24\cos I \sqrt{\eta_\star \sin I}}{
        2\pi \sin I + 12\sqrt{\eta_\star \sin I}\cos I
            + 2\pi \eta_\star \cos^2 I}.\label{eq:p_c_p2}
\end{equation}

To test this capture probability, we adopt a slightly different testing
methodology than in \autoref{s:p1}. For fixed initial condition $(\mu_0,
\phi_0)$ in zone I, we vary the initial $\eta \in [0.01, 0.02]$ and $\epsilon
\in 10^{[-3, -2]}$. In this way, every simulation experiences probabilistic
separatrix crossing at roughly comparable times. We then repeated this procedure
for three different values $I = 10^\circ, 20^\circ, 25^\circ$. A natural spread
in $\eta_\star$ was observed at each $I$ value, which permitted comparison to
\autoref{eq:p_c_p2} independently. The results for the three $I$ values are
presented in \autoref{fig:2_1hist}, where good agreement with
\autoref{eq:p_c_p2} is shown.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/1hist10.png}
    \end{subfigure}

    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/1hist20.png}
    \end{subfigure}

    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/2_toy2/1hist25.png}
    \end{subfigure}
    \caption{Simulations depicting $P_c(\eta_\star)$ at three different values
    $I = 10^\circ, 20^\circ, 25^\circ$. The top panel represents the binned
    outcomes as a function of $\eta_\star$, and the resulting estimates of $P_c$
    are compared with the analytical curve (red) from \autoref{eq:p_c_p2} in the
    bottom panel. Uncertainties in $P_c(\eta_\star)$ from data are estimated as
    $\sqrt{N}$ where $N$ is the count per bin. Also reported is the average
    value of $\eta_\star$.}\label{fig:2_1hist}
\end{figure}

\subsection{Conservation of Adiabatic Invariant}\label{ss:noncons}

Under the theory of adiabatic invariants (see e.g.\ \cite{henrard1982,
l_and_l}), trajectories following Hamiltonian equations of motion save for small
$\eta'$ conserve action variable $I$ except at separatrix crossing, where
\begin{equation}
    I \equiv \frac{1}{2\pi}\oint p\;\mathrm{d}q.
\end{equation}
$I$ has physical interpretation the enclosed phase space area of trajectories.
Its evolution can be seen in \autoref{fig:2_1testo4}. Under the assumption $I$
is conserved except at separatrix crossing, $\eta_\star$ is fixed implicitly by
$I = A_{sep}(\eta_\star)$ for $I$ the initial value of the action variable.
Since $A_{sep}$ is given by \autoref{eq:a_sep}, it would appear $\eta_\star$ can
be predicted exactly.

\textbf{NOTE:} The rest of this section is not very complete.

This turns out to be accurate in \autoref{ss:p2_spindown} but not in
\autoref{ss:p2_toy}, so it's probably non-physical/has to do with $\eta' = 0$
when $\mu = 0$. Not sure whether this is physically useful. Useful formulas for
$\rd{I}{t}$ can be taken from~\cite{henrard1982, l_and_l}:
\begin{equation}
    \rd{I}{\tau} = \s{\pd{H}{\eta}T(\eta) - \int\limits_0^{T(\eta)}
            \pd{H}{\eta}\;\mathrm{d}\tau} \eta'
        = -\pd{}{w}\p{\pd{S}{\eta}}\eta'.
\end{equation}
Here, $T(\eta)$ is the period of the orbit, $w$ is the angle variable conjugate
to $I$, and $S$ is the generating function for canonical transformation $\mu =
\pd{S}{\phi}, w = \pd{S}{I}$. These expressions are hard to evaluate, but maybe
they will tell us how much to expect $\rd{I}{\tau}$ to deviate from zero near
the separatrix?

@@

\subsection{Solution}

We consider now the full Cassini State system with under weak tidal friction
(see e.g.\ \cite{lai2012}). Note that generally $\abs{g}$ is a constant under
weak tidal friction, as it is an orbit-orbit coupling term, while $\alpha
\propto s$. Defining then $s_c$ critical spin such that $\frac{s}{s_c} \equiv
\frac{\alpha}{\abs{g}}$\footnote{$s_c$ has interpretation \emph{critical spin}
such that $\alpha, \abs{g}$ are equal.} and $\Omega_1$ the spin that $\hat{s}$
is coupled to, we can write down fully coupled evolution equations for $(\mu,
\phi, s)$:
\begin{subequations}\label{se:p3_eom}
    \begin{align}
        \pd{\phi}{\tau} &= \frac{s}{s_c}\mu
            - \p{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}} \cos \phi},\\
        \pd{\mu}{\tau} &= -\sin I \sin \phi +
            \epsilon \p{1 - \mu^2} \p{\frac{2\Omega_1}{s} - \mu},\\
        \rd{s}{\tau}
            &= \epsilon 2\Omega_1 \p{\mu - \frac{s}{2\Omega_1}\p{1 + \mu^2}}.
    \end{align}
\end{subequations}
The expressions for $H^{(0)}, H_4$ can also be written down
\begin{align}
    H(\mu, \phi; s) &= -\frac{s}{s_c}\frac{\mu^2}{2}
        + \mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi.\\
    H_4(s) &= -\sin I + \frac{s_c}{2s}\cos^2 I.
\end{align}

Note that here $s$ takes the role of adiabatically-varying parameter. Thus, when
computing $\Delta_{\pm}$ along the separatrix using \autoref{eq:delta}, there
are two contributions. In particular, we write
\begin{equation}
    \Delta_{\pm} = \int_{\mathcal{C}_{\pm}} \dot{\mu}^{(1)} +
        \frac{\dot{s}}{\dot{\phi}^{(0)}} \pd{h}{s}
        \;\mathrm{d}\phi.
\end{equation}
Both of these integrals were performed in preceeding sections, and we may
express in closed form (dropping some terms for simplicity)
\begin{align}
    \Delta_{\pm} \approx{}& \epsilon\z{
            \mp \frac{4\pi  \Omega_1}{s} + 8\sqrt{\frac{s_c}{s}\sin I}
            }\nonumber\\
        &+\epsilon\z{\frac{ \Omega_1 s_c}{s^2}
            \s{\pm 8\sqrt{\frac{s_c}{s} \sin I}
            \p{\frac{s}{2\Omega_1} - \frac{s_c\cos I}{s}}
                + 4\pi \frac{s_c}{s}\sin I}}.\label{eq:delta_p3}
\end{align}
We can then compute separatrix capture probability from zone III (which is $P_c
= \frac{\Delta_+ + \Delta_-}{\Delta_-}$), again evaluating at $s = s_\star$:
\begin{equation}
    P_c(s_\star) = \frac{16\sqrt{\frac{s_c}{s_\star}\sin I} +
        8\pi\frac{s_c}{s_\star}\sin I}{
            \frac{4\pi \Omega_1}{s_\star} - \frac{8\Omega_1
            s_c}{s_\star^2}\sqrt{\frac{s_c}{s_\star}\sin I}
            \p{\frac{s_\star}{2\Omega_1} - \frac{s_c \cos I}{s_\star}}}.
            \label{eq:p_c_p3}
\end{equation}
Note that for $\frac{s_c}{s_\star} < \eta_c$ that no separatrix (or resonance)
exists anymore, so we will generally assume $s_\star \gtrsim s_c, \Omega_1$.

To test \autoref{eq:p_c_p3}, we generate an evenly spaced grid of initial
conditions $(\mu_0, \phi_0)$ and initialize $s_0 = 10$. We then record $s_\star$
and the final ooutcome of the trajectory over many simulations. The agreement of
\autoref{eq:p_c_p3} with data can be observed in \autoref{fig:1_4stats}.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/1_weaktide/4_cross_hist0_05.png}
    \end{subfigure}

    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/1_weaktide/4_cross_hist0_20.png}
    \end{subfigure}

    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{../initial/1_weaktide/4_cross_hist0_40.png}
    \end{subfigure}

    \caption{Histogram of simulation and inferred $P_c(s_\star)$ compared with
    \autoref{eq:p_c_p3}. $s_c$ used are numbered at the top of each plot. $10^4$
    simulations were used for each plot, and a single $I = 20^\circ$ was
    used.}\label{fig:1_4stats}
\end{figure}

\subsection{Estimating $s_\star$}

In \autoref{s:p2}, the capture probability $P_c(\eta_\star)$ depended on
$\eta_\star$ which could be determined by conservation of the action varible
$I$. Such an argument is not possible in the weak tidal prescription as the
system is dissipitive thanks to $\dot{\mu}^{(1)}$ (see \autoref{se:p3_eom}).

However, instead we can write down $\rd{h}{\tau}$, just as \autoref{eq:dhdt}.
Since the initial value of $h$ can be computed from initial conditions, the ODE
can simply be integrated numerically to find $\tau_\star$ the separatrix
crossing time. In fact, for $\mu$ far from the separatrix, $\rd{h}{\tau} \approx
\frac{3s\mu^2}{2s_c}$, predicting
\begin{equation}
    \tau_\star = \frac{h(t = 0)}{3s\mu^2/2s_c} + C,
\end{equation}
where $C$ depends on the detairs of $\rd{h}{\tau}$ near the separatrix. This is
in good agreement with measurement (see top panel of \autoref{fig:1_4cross}).
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/4_cross0_20.png}
    \caption{TODO make this plot actually look reasonable. The top and bottom
    panels show $t_\star, s_\star$ as a function of $-h = H_4 - H$. The dotted
    red line in the top panel is $t_\star = \frac{h}{3s\mu^2/2s_c} + C$ where
    $C$ is a fitting parameter determined by the behavior of $\rd{h}{\tau}$ near
    the separatrix.}\label{fig:1_4cross}
\end{figure}

\subsection{High Misalignment as an Attractor}

Careful examination of \autoref{eq:delta_p3} reveals that for sufficiently large
$s_c$, $\Delta_+$ changes sign and becomes positive. When $\Delta_+ < 0$,
trajectories in zone I above the separatrix cannot enter the separatrix, as $h$
decreases over an orbit. However, if $\Delta_+ > 0$, then trajectories in zone I
can also enter the separatrix. The critical value of $s_c$ where $\Delta_+$
changes sign as a function of $s$ can thus be determined analytically. TODO
write down the formula and check it against my numerical solution earlier.

Furthermore, $\rd{\mu}{\tau} < 0$ for $\mu > \frac{2\Omega_1}{s}$, it is clear
that if $s$ is initially large, then aligned states $\mu > 0$ are driven towards
misalignment and the separatrix, where they may then be captured for
sufficiently large $s_c$. We find that $s_c \gtrsim 0.5 \Omega_1$ is sufficient
that an initial spin of $s_0 = 10$ will \emph{always} be captured into CS2. TODO
could shore this up a bit, find a good way to plot.

TODO find a clean plot to demonstrate this, I have the data already.
